---
title: "DATA608_module1_Exploratory Data Analysis"
author: "David Simbandumwe"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
editor_options: 
  chunk_output_type: inline
---


```{r setup}

knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
rm(list=ls())

```




**Principles of Data Visualization and Introduction to ggplot2**

I have provided you with data about the 5,000 fastest growing companies in the US, as compiled by Inc. magazine. lets read this in:

```{r}

inc <- read.csv("https://raw.githubusercontent.com/charleyferrari/CUNY_DATA_608/master/module1/Data/inc5000_data.csv", header= TRUE)

```

And lets preview this data:

```{r}

head(inc)

```



```{r}

summary(inc)

```

Think a bit on what these summaries mean. Use the space below to add some more relevant non-visual exploratory information you think helps you understand this data:

<b><i>
The skim() function outlines the structure of the dataset and key stats for the numeric and char variables. This is an interesting function because it deals with numeric and categorical data automatically. 
</i></b>

```{r}

library(skimr)

skim(inc)

```



<b><i>
The stat.desc() function computes additional descriptive statistics about the series in a data frame.  The function provides metrics on the number of na / null records and includes statistical measures such as std.dev, var etc.  
</i></b>

```{r}
library(pastecs)
library(dplyr)

stat_df <- stat.desc(Filter(is.numeric, inc))
stat_df <- format(stat_df, scientific = F, digits = 2, drop0trailing = TRUE)
stat_df

```


<b><i>
The ad.test() function tests for normal distribution for each variable in the dataset. With the resulting p-value of less than 0.05 we reject the null hypothesis for all variables. The variables are not normally distributed at a confidence level 0.95.  
</i></b>

```{r}

library(nortest)
ad_t <- ad.test(inc$Growth_Rate)
print(paste0(ad_t$method , ' variable ' , ad_t$data.name, ' normal distributed ', ad_t$p.value > 0.05 ))

ad_t <- ad.test(inc$Revenue)
print(paste0(ad_t$method , ' variable ' , ad_t$data.name, ' normal distributed ', ad_t$p.value > 0.05 ))

ad_t <- ad.test(inc$Employees)
print(paste0(ad_t$method , ' variable ' , ad_t$data.name, ' normal distributed ', ad_t$p.value > 0.05 ))


```


<b><i>
The rcorr() function creates a correlation matrix for numeric variables. It highlights the relationship between variable pairs in the dataset 
</i></b>


```{r}
library(Hmisc)

inc_num <- inc %>% dplyr::select(where(is.numeric))
rcorr(as.matrix(inc_num))

```

<b><i>
And finaly the average for variables for Revenue and Employees was computed using summarise in the dplyr package.
</i></b>

```{r}

inc %>% tidyr::drop_na(Revenue, Employees) %>% group_by(State) %>%
	select(City, State, Revenue, Employees) %>%
	summarise(avg_rev=mean(Revenue),avg_emp=mean(Employees)) %>% 
	arrange(desc(avg_rev))


inc %>% group_by(City) %>%
	select(City, State, Revenue, Employees) %>%
	summarise(avg_rev=mean(Revenue),avg_emp=mean(Employees)) %>% 
	arrange(desc(avg_rev))



```




## Question 1

Create a graph that shows the distribution of companies in the dataset by State (ie how many are in each state). There are a lot of States, so consider which axis you should use. This visualization is ultimately going to be consumed on a 'portrait' oriented screen (ie taller than wide), which should further guide your layout choices.

```{r, fig.height=4}

inc %>% group_by(State) %>% summarise(num = n()) %>%
	ggplot(aes(x=reorder(State,num), y=num, fill=State)) +
	coord_flip() + 
	theme_light() +
    geom_bar(stat = 'identity', width=0.3 , show.legend = FALSE) +
	labs(
	    x = 'State', 
	    y = 'Number of Companies', 
	    title = 'Distribution of Companies (by State)'
	)


```

## Quesiton 2

Lets dig in on the state with the 3rd most companies in the data set. Imagine you work for the state and are interested in how many people are employed by companies in different industries. Create a plot that shows the average and/or median employment by industry for companies in this state (only use cases with full data, use R's `complete.cases()` function.) In addition to this, your graph should show how variable the ranges are, and you should deal with outliers.


<b><i>
The state with the 3rd most companies is NY. The boxplot depicts the distribution of employees from each industry. The variable ranges for the data and captures the median and mean value of the number of Employees.
</i></b>



```{r}

st3 <- inc %>% group_by(State) %>% summarise(num = n()) %>%
				arrange(desc(num))

state_st <- st3$State[3]
print(paste0('State: ', state_st))
print(paste0('Original Record Count: ', st3$num[3]))
ny_df <- inc  %>% filter(State == state_st, complete.cases(.)) %>% arrange(Industry) %>% select(Name,Industry, Employees) 
print(paste0('Filter Complete Record: ', nrow(ny_df)))
ny_df <- ny_df %>% group_by(Industry) %>% filter(!(abs(Employees - median(Employees)) > 1.5*IQR(Employees)))
print(paste0('Filter Outliers: ', nrow(ny_df)))



ny_df %>% ggplot(aes(x=reorder(Industry,Employees), y=Employees)) +
	coord_flip() +
    geom_boxplot(show.legend = FALSE, outlier.colour = NA) +
	stat_summary(fun="mean", size=2, geom = "point", aes(color="Mean"))+
	stat_summary(fun="median", size=2, geom = "point", aes(color="Median"))+
	labs(
	    x = 'Industry', 
	    y = 'Number of Employees', 
	    title = 'NY Distribution of Employees (by Industry)'
	)


```





## Question 3

Now imagine you work for an investor and want to see which industries generate the most revenue per employee. Create a chart that makes this information clear. Once again, the distribution per industry should be shown.

<b><i>
The first step is to add the revenue per employee for each firm into the dataset. The distribution for each industry is depicted as box plot to highlight the population distribution. The graph includes the medium and the mean. 
</i></b>

```{r}

rev_df <- inc %>% mutate(rev_per_empl = Revenue/Employees)


print(paste0('Original Record Count: ', nrow(rev_df)))
rev_df <- rev_df  %>% filter(complete.cases(.)) %>% arrange(Industry) %>% select(Name, Industry, rev_per_empl, Revenue, Employees) 
print(paste0('Filter Complete Record: ', nrow(rev_df)))
rev_df <- rev_df %>% group_by(Industry) %>% filter(!(abs(rev_per_empl - median(rev_per_empl)) > 1.5*IQR(rev_per_empl)))
print(paste0('Filter Outliers: ', nrow(rev_df)))



rev_df %>% ggplot(aes(x=reorder(Industry,rev_per_empl), y=rev_per_empl)) +
	coord_flip() +
    geom_boxplot(show.legend = FALSE, outlier.colour = NA) +
	stat_summary(fun="mean", size=2, geom = "point", aes(color="Mean"))+
	stat_summary(fun="median", size=2, geom = "point", aes(color="Median"))+
	scale_y_continuous(labels = scales::comma) +
	labs(
	    x = 'Industry', 
	    y = 'Revenue per Employees', 
	    title = 'Distribution of Revenue Per Employee (by Industry)'
	)



```




