---
title: 'DATA608_module1-Interactive Visualizations'
author: 'David Simbandumwe'
output:
  html_document:
    toc: yes
    toc_float: yes
    theme: united
  pdf_document:
    toc: yes
editor_options:
  chunk_output_type: console
  markdown: 
    wrap: sentence
bibliography: references-hw3.bib
---





```{r setup}

knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
rm(list=ls())

```




**Principles of Data Visualization and Introduction to ggplot2**

I have provided you with data about the 5,000 fastest growing companies in the US, as compiled by Inc. magazine. lets read this in:

```{r}

#inc <- read.csv("https://raw.githubusercontent.com/charleyferrari/CUNY_DATA_608/master/module1/Data/inc5000_data.csv", header= TRUE)

```

And lets preview this data:

```{r}

#head(inc)

```



```{r}

#summary(inc)

```

Think a bit on what these summaries mean. Use the space below to add some more relevant non-visual exploratory information you think helps you understand this data:

<b><i>
The skim() function outlines the structure of the dataset and key stats for the numeric and char variables. This is an interesting function because it deals with numeric and categorical data automatically. 
</i></b>

```{r}

library(skimr)
#skim(inc)

```






